---
title: "Exploratory Data Analysis"
output: 
  html_notebook:
    toc: true
    toc_float: true
---

```{r setup, message = FALSE}
library(tidyverse)
library(jsonlite)

report_theme <- ggplot2::theme_bw() +
  ggplot2::theme(panel.border = ggplot2::element_rect(color = "black",
                                                      fill = NA),
                 panel.grid.major = ggplot2::element_line(color = "black",
                                                          linetype = "dotted"),
                 panel.grid.minor = ggplot2::element_line(color = "black",
                                                          linetype = "dotted"))
```

In this notebook I will be doing an exploratory data analysis looking to answer 
the question: 

> What factors correlate with users booking?

Particularly, I'm interesting in looking at what factors correlate with the 
conversion of leads to bookings. As such, I will primarily work with the leads 
data set with relevant information from the referral data set and a logical 
column indicating a booking event.

# Data Import and Preparation

First, I need to import and prepare the data for analysis.

```{r, message = FALSE}
room_key_data <- read_tsv("data.tsv", progress = FALSE) %>% 
  rownames_to_column() %>% 
  rename(key = rowname) %>% 
  mutate(lead_event = !(lead == "null"), 
         booking_event = !(booking == "null"))
```

```{r}
referral_key <- room_key_data %>% 
  filter(referral != "null") %>% 
  select(key, lead_event, booking_event)

referral <- room_key_data %>% 
  filter(referral != "null") %>%
  .$referral %>% 
  map_df(fromJSON) %>% 
  bind_cols(referral_key) %>% 
  select(key, everything()) %>% 
  rename(visitor_id = `visitor-id`, 
         event_id_referral = `event-id`, 
         ts_referral = ts,
         location_tid = `location-tid`, 
         partner_code = `partner-code`, 
         check_in = `check-in`, 
         check_out = `check-out`)
```

```{r}
lead_key <- room_key_data %>% 
  filter(lead != "null") %>% 
  select(key, booking_event)

lead <- room_key_data %>% 
  filter(lead != "null") %>%
  .$lead %>% 
  map_df(fromJSON) %>% 
  bind_cols(lead_key) %>% 
  select(key, everything()) %>% 
  rename(visitor_id = `visitor-id`, 
         event_id_lead = `event-id`, 
         ts_lead = ts, 
         rate_partner = `rate-partner`)
```

```{r}
booking_key <- room_key_data %>% 
  filter(booking != "null") %>% 
  select(key)

booking <- room_key_data %>% 
  filter(booking != "null") %>%
  .$booking %>%  
  map_df(fromJSON) %>% 
  bind_cols(booking_key) %>% 
  select(key, everything())
```

## Preparation

Here I want to combine some information from the referral data set into the 
lead data set. As part of this, I also rename a few columns and convert 
some of the columns to different data types. Specifically the ts columns are 
converted to datetime and the check_in and check_out columns are converted 
to dates.

```{r}
referral_info <- referral %>% 
  filter(lead_event == TRUE) %>% 
  select(-event_id_referral, -type, -lead_event, -booking_event)

lead_ref <- lead %>% 
  select(-event_id_lead, - type) %>% 
  left_join(referral_info, by = c("key", "visitor_id")) %>% 
  rename(rate_partner_lead = rate_partner, 
         partner_code_referral = partner_code) %>% 
  mutate(ts_lead = lubridate::as_datetime(ts_lead / 1000), 
         ts_referral = lubridate::as_datetime(ts_referral / 1000), 
         check_in = lubridate::ymd(check_in), 
         check_out = lubridate::ymd(check_out))
```

```{r}
lead_ref
```

# Analysis

The following are some initial questions looking at how some of the variables 
interact. 

## Time Difference 

I want to look at whether or not the time difference between lead and referral 
corresponds with a booking.

```{r}
lead_ref %>% 
  mutate(ts_time_diff = as.numeric(ts_lead - ts_referral)) %>% 
  # filter(ts_time_diff < 3000) %>%
  ggplot(aes(x = booking_event, y = ts_time_diff)) + 
  geom_boxplot()
```

```{r}
lead_ref %>% 
  mutate(ts_time_diff = as.numeric(ts_lead - ts_referral)) %>% 
  filter(ts_time_diff < 3000) %>%
  ggplot(aes(x = booking_event, y = ts_time_diff)) + 
  geom_boxplot()
```

```{r}
lead_ref %>% 
  mutate(ts_time_diff = as.numeric(ts_lead - ts_referral)) %>% 
  filter(ts_time_diff < 500) %>%
  ggplot(aes(x = ts_time_diff, fill = booking_event)) + 
  geom_histogram(binwidth = 10) + 
  facet_grid(booking_event ~ ., scales = "free_y")
```

There does not appear to be a large difference between the times at which 
booking events occur or don't occur. There is possibly a greater concentation 
of booking events at short times (time difference < 100 s), but both display a 
similar tail at large time differences. 

## Check-in Time Difference

Here I want to look at the time difference between the proposed check-in date 
and both the referral and lead time stamp date.

First, I want to look at the presence of any NAs in the date differences. 

```{r}
lead_ref %>% 
  mutate(ts_time_diff = as.numeric(check_in - lubridate::as_date(ts_referral))) %>%
  group_by(booking_event) %>% 
  summarize(check_in_NA = sum(is.na(check_in)), 
            ts_referral_NA = sum(is.na(ts_referral)), 
            time_diff_NA = sum(is.na(ts_time_diff)))
```

Here I can see that there are a number of NA's in the check-in data and a few
in the timestamp from the referral data. There were no NA's in the timestamp 
for the lead data. Since the check-in NA number and the time difference NA 
number are the same, I can assume that the NA's from the referral data overlap 
with those from the check-in data. 

For this next analysis, I will remove all of the NA's as they cannot contribute.

### Check-in - Referral 

First, I will look at the difference in time between the check-in date and the 
referral timestamp.

```{r}
lead_ref %>% 
  filter(!is.na(check_in)) %>% 
  mutate(ts_time_diff = as.numeric(check_in - lubridate::as_date(ts_referral))) %>%
  ggplot(aes(x = ts_time_diff, fill = booking_event)) + 
  geom_histogram(binwidth = 1) + 
  facet_grid(booking_event ~ ., scales = "free_y") +
  coord_cartesian(xlim = c(-10, 100)) +
  labs(x = "Check-in - Referral Date (days)", 
       y = "Count", 
       fill = "Booking Event")
```

```{r}
lead_ref %>% 
  filter(!is.na(check_in)) %>% 
  mutate(ts_time_diff = as.numeric(check_in - lubridate::as_date(ts_referral))) %>%
  ggplot(aes(x = booking_event, y = ts_time_diff)) + 
  geom_boxplot() +
  coord_cartesian(ylim = c(-10, 100)) +
  labs(y = "Check-in - Referral Date (days)", 
       x = "Booking Event")
```

Both the histogram and the boxplot show that the majority of the events occur 
within 7 days of the referral. According to the boxplot, the median time 
difference for a TRUE booking event is slightly lower than a FALSE booking 
event. The occurance of a TRUE booking event is also slightly more concentrated 
compared to a FALSE booking event. 

### Check-in - Lead

```{r}
lead_ref %>% 
  filter(!is.na(check_in)) %>% 
  mutate(ts_time_diff = as.numeric(check_in - lubridate::as_date(ts_lead))) %>%
  ggplot(aes(x = ts_time_diff, fill = booking_event)) + 
  geom_histogram(binwidth = 1) + 
  facet_grid(booking_event ~ ., scales = "free_y") +
  coord_cartesian(xlim = c(-10, 100)) +
  labs(x = "Check-in - Lead Date (days)", 
       y = "Count", 
       fill = "Booking Event")
```

```{r}
lead_ref %>% 
  filter(!is.na(check_in)) %>% 
  mutate(ts_time_diff = as.numeric(check_in - lubridate::as_date(ts_lead))) %>%
  ggplot(aes(x = booking_event, y = ts_time_diff)) + 
  geom_boxplot() +
  coord_cartesian(ylim = c(-10, 100)) +
  labs(y = "Check-in - Lead Date (days)", 
       x = "Booking Event")
```

The check-in lead time difference shows a similar pattern to the check-in 
referral time difference. 

Of note, there are a few interesting data points in this set. There are a small
number of points with check-in times earlier than the time stamp for the 
referral or lead. When combined with a TRUE booking event, this indicates that 
these bookings were done after the time. Most likely, these are some typos in 
the check-in date.

```{r}
lead_ref %>% 
  mutate(ts_time_diff = as.numeric(check_in - lubridate::as_date(ts_lead))) %>%
  filter(booking_event == TRUE) %>% 
  filter(ts_time_diff < 0)
```

## Partner 

For this section, I want to look at whether the specific partner corresponds 
with an increased booking rate. In this section I will look at the rate partner 
that the lead goes to, the referral partner that the user came from, and whether 
those partners are the same from referral to lead.

### Rate Partner Lead Data

The lead rate partners are shown using a bar plot with the existence of a 
booking event given by the colors. The first bar plot gives a simple count of 
the number of TRUE and FALSE booking events for each partner. The second plot 
gives the proportion of booking events for each partner.

```{r}
lead_ref %>% 
  ggplot(aes(x = rate_partner_lead, fill = booking_event)) +
  geom_bar() +
  coord_flip() +
  labs(y = "Count", 
       x = "Lead Rate Partner", 
       fill = "Booking Event")
```

```{r}
lead_ref %>% 
  ggplot(aes(x = rate_partner_lead, fill = booking_event)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(y = "Proportion", 
       x = "Lead Rate Partner", 
       fill = "Booking Event")
```

The proportion bar plot shows that the partners PEGS, MRT, IHG, HLTN, and CHC 
have the highest proportion of booking events. MRT, IGH< HLTN, and CHC are also 
the partners with the largest number of leads which  may influence their higher 
rate of booking.

### Referral Partner Data

The referral partner data is given in two bar plots as with the lead data. The 
first bar plot is acount of the number of TRUE or FALSE booking events for 
each referral partner. The second bar plot is the propotion of TRUE and FALSE 
booking events for each referral partner.

```{r}
lead_ref %>% 
  ggplot(aes(x = partner_code_referral, fill = booking_event)) +
  geom_bar() +
  coord_flip() +
  labs(y = "Count", 
       x = "Referral Partner", 
       fill = "Booking Event")
```

```{r}
lead_ref %>% 
  ggplot(aes(x = partner_code_referral, fill = booking_event)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(y = "Proportion", 
       x = "Referral Partner", 
       fill = "Booking Event")
```

Here, the proportion data shows that MRT, IHG, and HLTN are the three biggest 
sources of booking events. However, more of the referral partners have 
noticeable booking event proportions.

### Same Partner Data

For the same partner data I am interested in calculating which users followed a 
lead from the same partner that referred them to Room Key. For this data, I 
noticed that there are a few referral partners that show up as NA. Since these 
are only a small number of the total leads and bookings, I will filter these 
out.

```{r}
lead_ref %>% 
  summarize(sum(is.na(partner_code_referral)))
```

As with the previous partner sections, I will plot this data in two ways. First 
a simple count of TRUE or FALSE booking events for partners that are the same 
from referral to lead and for ones that are different. Second a bar plot of the 
proportions of booking events.

```{r}
lead_ref %>% 
  mutate(same_partner = (partner_code_referral == rate_partner_lead)) %>% 
  filter(!is.na(same_partner)) %>% 
  ggplot(aes(x = same_partner, fill = booking_event)) +
  geom_bar() +
  coord_flip() +
  labs(y = "Count", 
       x = "Same Partner", 
       fill = "Booking Event")
```

```{r}
lead_ref %>% 
  mutate(same_partner = (partner_code_referral == rate_partner_lead)) %>% 
  filter(!is.na(same_partner)) %>% 
  ggplot(aes(x = same_partner, fill = booking_event)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(y = "Proportion", 
       x = "Same Partner", 
       fill = "Booking Event")
```

Interestingly, the vast majority of booking events occurs when the partners are 
the same from referral to lead. This is true, even though close to 36% of the 
leads occur when the partners are not the same. Given that, only 12% of the 
bookings occur when partners are not the same.

```{r}
lead_ref %>% 
  mutate(same_partner = (partner_code_referral == rate_partner_lead)) %>% 
  filter(!is.na(same_partner)) %>% 
  group_by(booking_event) %>% 
  summarize(prop_same = mean(same_partner), 
            prop_diff = 1 - prop_same)
```

```{r}
lead_ref %>% 
  mutate(same_partner = (partner_code_referral == rate_partner_lead)) %>% 
  filter(!is.na(same_partner)) %>% 
  # group_by(booking_event) %>% 
  summarize(prop_same = mean(same_partner), 
            prop_diff = 1 - prop_same)
```

## Location-tid and udicode

```{r}
lead_ref %>% 
  group_by(location_tid) %>% 
  summarize(total = n(), 
            total_true = sum(booking_event), 
            total_false = total - total_true, 
            prop_true = total_true / total, 
            prop_false = total_false / total) %>% 
  arrange(desc(prop_true))
```

```{r}
lead_ref %>% 
  group_by(udicode) %>% 
  summarize(total = n(), 
            total_true = sum(booking_event), 
            total_false = total - total_true, 
            prop_true = total_true / total, 
            prop_false = total_false / total) %>% 
  filter(total_true > 1) %>% 
  arrange(desc(prop_true))
```

