---
title: "Exploratory Data Analysis"
output: 
  html_notebook:
    toc: true
    toc_float: true
---

```{r setup, message = FALSE}
library(tidyverse)
library(jsonlite)
```

The two questions that I have been asked to answer to the Room Key Data 
Science Challenge are: 

* What factors correlate with users booking? 
* What can Room Key do to increase booking conversion rate?

To begin, I want to do some exploratory data analysis to get a sense for the 
data.

# Data Import

The first step is to import the data to R. The data is stored as a tab-separated 
value format with three columns. The elements of each column contain an entry 
for referrals, leads, and booking in JSON format.

```{r, message = FALSE}
room_key_data <- read_tsv("data.tsv", progress = FALSE)
```

With the data imported, I want to take a peak at what it looks like. 

```{r}
room_key_data
```

In this form the data is pretty difficult to parse. I want to take each column 
of the `room_key_data` object (referral, lead, and booking) and split them into 
individual data sets. These individual data sets can then be parsed by their 
JSON format and turned into a data frame. 

# Data Cleaning

The first step of the data cleaning will be to split each data set apart. For 
each of the data sets, I will filter out the null entries as these are ones 
that did not lead to any leads or booking. The filtering step will also occur 
for the referral set, but it should not lead to the removal of any entries. 
The `fromJSON` function will then be applied to every element of the column 
and a new data frame will be formed with all of the information spread out 
for each data set.

## Referral

The first set to look at is the referral set. 