---
title: "Exploratory Data Analysis"
output: 
  html_notebook:
    toc: true
    toc_float: true
---

```{r setup, message = FALSE}
library(tidyverse)
library(jsonlite)

report_theme <- ggplot2::theme_bw() +
  ggplot2::theme(panel.border = ggplot2::element_rect(color = "black",
                                                      fill = NA),
                 panel.grid.major = ggplot2::element_line(color = "black",
                                                          linetype = "dotted"),
                 panel.grid.minor = ggplot2::element_line(color = "black",
                                                          linetype = "dotted"))
```

In this notebook I will be doing an exploratory data analysis looking to answer 
the question: 

> What factors correlate with users booking?

Particularly, I'm interesting in looking at what factors correlate with the 
conversion of leads to bookings. As such, I will primarily work with the leads 
data set with relevant information from the referral data set and a logical 
column indicating a booking event.

# Data Import and Preparation

First, I need to import and prepare the data for analysis.

```{r, message = FALSE}
room_key_data <- read_tsv("data.tsv", progress = FALSE) %>% 
  rownames_to_column() %>% 
  rename(key = rowname) %>% 
  mutate(lead_event = !(lead == "null"), 
         booking_event = !(booking == "null"))
```

```{r}
referral_key <- room_key_data %>% 
  filter(referral != "null") %>% 
  select(key, lead_event, booking_event)

referral <- room_key_data %>% 
  filter(referral != "null") %>%
  .$referral %>% 
  map_df(fromJSON) %>% 
  bind_cols(referral_key) %>% 
  select(key, everything()) %>% 
  rename(visitor_id = `visitor-id`, 
         event_id_referral = `event-id`, 
         ts_referral = ts,
         location_tid = `location-tid`, 
         partner_code = `partner-code`, 
         check_in = `check-in`, 
         check_out = `check-out`)
```

```{r}
lead_key <- room_key_data %>% 
  filter(lead != "null") %>% 
  select(key, booking_event)

lead <- room_key_data %>% 
  filter(lead != "null") %>%
  .$lead %>% 
  map_df(fromJSON) %>% 
  bind_cols(lead_key) %>% 
  select(key, everything()) %>% 
  rename(visitor_id = `visitor-id`, 
         event_id_lead = `event-id`, 
         ts_lead = ts, 
         rate_partner = `rate-partner`)
```

```{r}
booking_key <- room_key_data %>% 
  filter(booking != "null") %>% 
  select(key)

booking <- room_key_data %>% 
  filter(booking != "null") %>%
  .$booking %>%  
  map_df(fromJSON) %>% 
  bind_cols(booking_key) %>% 
  select(key, everything())
```

## Preparation

Here I want to combine some information from the referral data set into the 
lead data set. As part of this, I also rename a few columns and convert 
some of the columns to different data types. Specifically the ts columns are 
converted to datetime and the check_in and check_out columns are converted 
to dates.

```{r}
referral_info <- referral %>% 
  filter(lead_event == TRUE) %>% 
  select(-event_id_referral, -type, -lead_event, -booking_event)

lead_ref <- lead %>% 
  select(-event_id_lead, - type) %>% 
  left_join(referral_info, by = c("key", "visitor_id")) %>% 
  rename(rate_partner_lead = rate_partner, 
         partner_code_referral = partner_code) %>% 
  mutate(ts_lead = lubridate::as_datetime(ts_lead / 1000), 
         ts_referral = lubridate::as_datetime(ts_referral / 1000), 
         check_in = lubridate::ymd(check_in), 
         check_out = lubridate::ymd(check_out))
```

```{r}
lead_ref
```

# Analysis

The following are some initial questions looking at how some of the variables 
interact. 

## Time Difference 

I want to look at whether or not the time difference between lead and referral 
corresponds with a booking.

```{r}
lead_ref %>% 
  mutate(ts_time_diff = ts_lead - ts_referral)
```

