---
title: "Exploratory Data Analysis"
output: 
  html_notebook:
    toc: true
    toc_float: true
---

```{r setup, message = FALSE}
library(tidyverse)
library(jsonlite)
```

The two questions that I have been asked to answer to the Room Key Data 
Science Challenge are: 

* What factors correlate with users booking? 
* What can Room Key do to increase booking conversion rate?

To begin, I want to do some exploratory data analysis to get a sense for the 
data.

# Data Import

The first step is to import the data to R. The data is stored as a tab-separated 
value format with three columns. The elements of each column contain an entry 
for referrals, leads, and booking in JSON format.

```{r, message = FALSE}
room_key_data <- read_tsv("data.tsv", progress = FALSE)
```

With the data imported, I want to take a peak at what it looks like. 

```{r}
room_key_data
```

In this form the data is pretty difficult to parse. I want to take each column 
of the `room_key_data` object (referral, lead, and booking) and split them into 
individual data sets. These individual data sets can then be parsed by their 
JSON format and turned into a data frame. 

# Data Cleaning

The first step of the data cleaning will be to split each data set apart. For 
each of the data sets, I will filter out the null entries as these are ones 
that did not lead to any leads or booking. The filtering step will also occur 
for the referral set, but it should not lead to the removal of any entries. 
The `fromJSON` function will then be applied to every element of the column 
and a new data frame will be formed with all of the information spread out 
for each data set.

## Referral

The first set to look at is the referral set. 

```{r}
referral <- room_key_data %>% 
  filter(referral != "null") %>%
  .$referral %>% 
  map_df(fromJSON)
```

Note that converting this into a data frame is time intensive. Here is a peak 
at the completed data frame. 

```{r}
referral
```

## Lead

Next I want to convert the lead column into a data frame.

```{r}
lead <- room_key_data %>% 
  filter(lead != "null") %>%
  .$lead %>% 
  map_df(fromJSON)
```

Here is the lead data set. 

```{r}
lead
```

## Booking

Lastly, I want to convert the booking data set into a data frame. 

```{r}
booking <- room_key_data %>% 
  filter(booking != "null") %>%
  .$booking %>% 
  map_df(fromJSON)
```

And here is the booking data set.

```{r}
booking
```

# Data Details

Now I want to explore the details of each data set. What kind of variables 
am I working with in each one? 

## Referral 

In the referral data set there are 9 columns. 

* visitor-id 
* event-id
* ts 
* type 
* location-tid 
* partner-code 
* check-in 
* check-out 
* udicode

```{r}
room_key_data %>% 
  filter(booking != "null") %>% 
  rownames_to_column() %>% 
  filter(rowname == 1) %>% 
  select(-rowname)
```

Doing a visual inspection of an observation that contains a referral, lead, 
and booking, it appears that the `visitor-id` variable is a unique identifier 
for a user/visitor. The `event-id` variable however is not a unique identifier 
of an observation within, but rather an identifier for an individual referral, 
lead, or booking event. That is to say, for a single row there are different 
event-id's for each referral, lead, and booking event.

To get a sense of what each column contains I am going to do a few simple 
explorations. First, the total number of observations.

```{r}
referral %>% 
  count()
```

### event-id

Next, the number of distinct event-id's.

```{r}
referral %>% 
  distinct(`event-id`) %>% 
  count()
```

The number of distinct event-id's and the total number of observations are 
equal, indicating that the `event-id` column does truly represent each unique 
referral event. 

### visitor-id

Next, I want to look at the `visitor-id` column.

```{r}
referral %>% 
  distinct(`visitor-id`) %>% 
  count()
```

Here, the number of unique visitor id's is less than the total number of 
observations. This indicates that there are a few return visitors to the site.

### ts

Next, I want to understand what the `ts` column is. It appears to be a large 
number and may be a timestamp written in Unix time. 

```{r}
referral %>% 
  select(ts) %>% 
  mutate(timestamp = lubridate::as_datetime(ts))
```

The timestamps produced assuming that `ts` is in Unix time of seconds produces 
timestamps in the year 49757 which is obviously incorrect. If I assume that the 
`ts` column is actually in units of milliseconds then I get the following 
timestamps.

```{r}
referral %>% 
  select(ts) %>% 
  mutate(timestamp = lubridate::as_datetime(ts / 1000)) %>% 
  summarize(min_date = min(timestamp), max_date = max(timestamp))
```

These timestamps match the date range given in the challenge prompt.

### type

Next I want to determine how many `type` show up in the set. 

```{r}
referral %>% 
  distinct(type)
```

There is only one type that appears in this data set, so it is likely that 
the type can be discarded unless I plan to combine the three data sets together. 

### location-tid

Next is the `location-tid` column.

```{r}
referral %>% 
  distinct(`location-tid`)
```

There are about 13k distinct elements in `location-tid`. How many times does 
each distinct element appear? 

```{r}
referral %>% 
  select(`location-tid`) %>% 
  group_by(`location-tid`) %>% 
  count() %>% 
  arrange(desc(n))
```

It looks like the largest appearance is made by NA. Additional, there appear 
to be many repeat appearances of these elements. This indicates that this may 
be a variable I can examine deeper. Assuming I can determine the origin of the 
NA elements.

### Partner Code

Now, I want to look at the `partner-code` and see how many unique partners 
there are. 

```{r}
referral %>% 
  distinct(`partner-code`)
```

There are 7 unique partners that appear with the following frequency. 

```{r}
referral %>% 
  group_by(`partner-code`) %>% 
  count() %>% 
  arrange(desc(n))
```

### check-in and check-out 

Next, I want to examine the `check-in` and `check-out` columns. These are 
fairly self-explantory in that they are the dates the visitor wants to check in 
and check out for their trip. These variable may be of interest as either a way 
of looking at correlations between booking and check in date, correlations 
between book and length of trip, or correlations between booking and time 
between referral date and check-in date.

```{r}
referral %>% 
  select(`check-in`, `check-out`) %>% 
  mutate(check_in = lubridate::ymd(`check-in`), 
         check_out = lubridate::ymd(`check-out`))
```

### udicode

Next, I look at the `udicode`. 

```{r}
referral %>% 
  group_by(udicode) %>% 
  count() %>% 
  arrange(desc(n))
```

This is another variable that has a number of repeated values. This could be 
another variable to examine for correlations, though the vast majority of 
this variable is encoded as NA. 

## Lead 

In the `lead` data set there are 5 columns: 

* visitor-id
* event-id 
* ts 
* type 
* rate-partner

First, I want to verify that all of the columns similar to the same named 
columns in referral are truly similar. 

```{r}
lead %>% 
  count()
```

```{r}
lead %>% 
  distinct(`event-id`) %>% 
  count()
```

As before, `event-id` is the number of unique lead events. 

```{r}
lead %>% 
  distinct(`visitor-id`) %>% 
  count()
```

Also as before, the `visitor-id` variable is unique for each visitor, but there 
are likely a few repeat visitors.

```{r}
lead %>% 
  select(ts) %>% 
  mutate(timestamp = lubridate::as_datetime(ts / 1000)) %>% 
  summarize(min_date = min(timestamp), max_date = max(timestamp))
```

Here, `ts` appears to be the timestamp for when the lead event occured. 

```{r}
lead %>% 
  distinct(type)
```

As with the referral set, the type for lead only refers to the api used.

### rate-partner

The only new variable in this set is `rate-partner`. 

```{r}
lead %>% 
  group_by(`rate-partner`) %>% 
  count() %>% 
  arrange(desc(n))
```

Like the `partner-code` variable from the referral data set, the `rate-partner` 
refers to the hotel group that the lead linked to.

## Booking

In the booking data set there are 11 columns: 

* currency
* nights 
* visitor-id 
* check-in
* type 
* check-out 
* rate-partner 
* ts 
* rooms 
* event-id 
* total-amount

Some of these variables overlap with the variables from the previous sets. I 
first want to look at those variables and verify they respond similarly.

First, I want to verify that all of the columns similar to the same named 
columns in referral are truly similar. 

```{r}
booking %>% 
  count()
```

```{r}
booking %>% 
  distinct(`event-id`) %>% 
  count()
```

As before, `event-id` is the number of unique booking events. 

```{r}
booking %>% 
  distinct(`visitor-id`) %>% 
  count()
```

Also as before, the `visitor-id` variable is unique for each visitor, but there 
are likely a few repeat visitors. For the booking data set, there are almost an 
equal number of visitors as unique events indicating that there are few repeat 
bookings from the same visitor in the time span recorded.

```{r}
booking %>% 
  select(ts) %>% 
  mutate(timestamp = lubridate::as_datetime(ts / 1000)) %>% 
  summarize(min_date = min(timestamp), max_date = max(timestamp))
```

Here, `ts` appears to be the timestamp for when the booking event occured. 

```{r}
booking %>% 
  distinct(type)
```

As with the referral set, the type for booking only refers to the api used.

```{r}
booking %>% 
  select(`check-in`, `check-out`) %>%
  mutate(check_in = lubridate::ymd(`check-in`), 
         check_out = lubridate::ymd(`check-out`))
```

While looking at the check-in and check-out dates for the booking, there are 
four dates that apparently failed to parse. A further investigation of these 
dates showed that there are 10 check-out dates filled in with NA. Additionally, 
there are several dates, the four not parsed, that are in non-standard date 
formats.

```{r}
booking %>% 
  filter(is.na(`check-out`)) %>% 
  mutate(check_in = lubridate::ymd(`check-in`), 
         check_out = lubridate::ymd(`check-out`))
```

The four check-in dates that can't be parsed also have non-standard entries in 
some of the other elements. All of these come from the rate-partner CHC. 

```{r}
booking %>% 
  filter(`rate-partner` == "CHC")
```

Note that not all of the CHC bookings have the non-standard entry for nights or 
check-ins.

```{r}
booking %>% 
  group_by(`rate-partner`) %>% 
  count() %>% 
  arrange(desc(n))
```

The `rate-partner` entries are similar to that in the lead data set. 

### currency

```{r}
booking %>% 
  distinct(currency)
```

The `currency` column contains what type of currency was used to pay for the 
booking. 

### nights

```{r}
booking %>% 
  distinct(nights)
```

The `nights` column is the number of nights a booking was made for. Note that 
there are some non-standard entries for the nights as discussed in the 

### rooms 

```{r}
booking %>% 
  group_by(rooms) %>% 
  count() %>% 
  arrange(desc(n))
```

Again, there are a few non-standard entries here including a few NA entries. 
These are likely from a hotel that did not report their rooms.

### total-amount

```{r}
booking %>% 
  select(currency, `total-amount`)
```

The `total-amount` column is simply the total amount of money paid for the 
booking. Note that there are a few NA entries.

# Data Import and Cleaning Procedure

The following is a procedure to be used in the rest of the project for the 
import, organizing, and cleaning of the data. This will include creating a 
key for each data set for appropriate joining and the creation of a few new 
variables.

```{r, message = FALSE}
room_key_data <- read_tsv("data.tsv", progress = FALSE) %>% 
  rownames_to_column() %>% 
  rename(key = rowname) %>% 
  mutate(lead_event = !(lead == "null"), 
         booking_event = !(booking == "null"))
```

```{r}
referral_key <- room_key_data %>% 
  filter(referral != "null") %>% 
  select(key, lead_event, booking_event)

referral <- room_key_data %>% 
  filter(referral != "null") %>%
  .$referral %>% 
  map_df(fromJSON) %>% 
  bind_cols(referral_key) %>% 
  select(key, everything())
```

```{r}
lead_key <- room_key_data %>% 
  filter(lead != "null") %>% 
  select(key)

lead <- room_key_data %>% 
  filter(lead != "null") %>%
  .$lead %>% 
  map_df(fromJSON) %>% 
  bind_cols(lead_key) %>% 
  select(key, everything())
```

```{r}
booking_key <- room_key_data %>% 
  filter(booking != "null") %>% 
  select(key)

booking <- room_key_data %>% 
  filter(booking != "null") %>%
  .$booking %>%  
  map_df(fromJSON) %>% 
  bind_cols(booking_key) %>% 
  select(key, everything())
```
