---
title: "Exploratory Data Analysis"
output: 
  html_notebook:
    toc: true
    toc_float: true
---

```{r setup, message = FALSE}
library(tidyverse)
library(jsonlite)
```

The two questions that I have been asked to answer to the Room Key Data 
Science Challenge are: 

* What factors correlate with users booking? 
* What can Room Key do to increase booking conversion rate?

To begin, I want to do some exploratory data analysis to get a sense for the 
data.

# Data Import

The first step is to import the data to R. The data is stored as a tab-separated 
value format with three columns. The elements of each column contain an entry 
for referrals, leads, and booking in JSON format.

```{r, message = FALSE}
room_key_data <- read_tsv("data.tsv", progress = FALSE)
```

With the data imported, I want to take a peak at what it looks like. 

```{r}
room_key_data
```

In this form the data is pretty difficult to parse. I want to take each column 
of the `room_key_data` object (referral, lead, and booking) and split them into 
individual data sets. These individual data sets can then be parsed by their 
JSON format and turned into a data frame. 

